---
title: "Class 11: Genome Informatics"
author: "Anh Tran"
format: pdf
---

```{r}
mxl <- read.csv("MXL.csv", row.names=1)
```

```{r}
View(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl)
```

#Section 4: Population Analysis

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes. 

```{r}
url <- "https://bioboot.github.io/bimm143_F22/class-material/rs8067378_ENSG00000172057.6.txt"

file <- read.table(url, row.names=1)
head(file)
```

The sample size for each genotype

```{r}
table(file$geno)
```

```{r}
A <- file[file$geno=="A/A",]
B <- file[file$geno=="A/G",]
C <- file[file$geno=="G/G",]
```

The median values for A/A, A/G, and G/G correspondingly are:

```{r}
median(A$exp)
median(B$exp)
median(C$exp)
```

>Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

```{r}
library(ggplot2)
```

```{r}
plot <- ggplot(file) + aes(geno=geno, y=exp, fill=geno) + geom_boxplot(notch=TRUE)
plot
```

The expression levels are lower for G/G and higher for A/A meaning that having G/G is associated with reduced expression of this gene. Yes, SNP affects the expression of ORMDL3.
