---
title: "Class 10: Halloween Mini-Project"
author: "Anh Tran"
format: pdf
---

```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
candy <- read.csv(url, row.names=1)
head(candy)
```

>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

My favorite candy is Milky Way

```{r}
#rownames(candy)
candy["Milky Way",]$winpercent
```

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

```{r}
#install.packages("skimr")
```

```{r}
#library("skimr")
#Instead of using library() to load the whole package, can use this function instead for just one function

skimr::skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, winpercent is the only column that looks different from other other columns as its values are much higher than 0. 

>Q7. What do you think a zero and one represent for the candy$chocolate column?

If contains chocolate or not (1=yes, 0=no)

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

```{r}
#install.packages("ggplot2")
```

```{r}
library(ggplot2)
ggplot(candy) + 
  aes(winpercent) +
  geom_histogram(bins=10, fill="black", col="pink")
```

>Q9. Is the distribution of winpercent values symmetrical?

No the distribution is not symmetrical. It is slightly skewed to the left. 

>Q10. Is the center of the distribution above or below 50%?

```{r}
median(candy$winpercent)
```

It is below 50% 

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate.inds <- as.logical(candy$chocolate)
chocolate.wins <- candy[chocolate.inds,]$winpercent
mean(chocolate.wins)

fruity.inds <- as.logical(candy$fruity)
fruity.wins <- candy[fruity.inds,]$winpercent
mean(fruity.wins)
```

```{r}
t.test(chocolate.wins, fruity.wins)
```

Chocolate candy is higher ranked than fruit candy

>Q12. Is this difference statistically significant?

Yes, this difference is statistically significant

## 3. Overall Candy Rankings

>13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(-candy$winpercent),], n=5)
```

>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)
ggplot(candy) + 
  aes(winpercent, rownames(candy)) + 
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) + 
  geom_col() +
  coord_fixed()
```


```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] ="chocolate"
my_cols[as.logical(candy$bar)]="brown"
my_cols[as.logical(candy$fruity)]="red"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) + 
  geom_col(fill=my_cols)
```

```{r}
#Save the last ggplot as image
ggsave("tmp.png")
```

>Q17. What is the worst ranked chocolate candy?

Sixlets

>Q18. What is the best ranked fruity candy?

Starburst

## 4. Taking a look at pricepercent

```{r}
#install.packages("ggrepel")
library(ggrepel)

ggplot(candy) + 
  aes(winpercent, pricepercent, label=rownames(candy)) + 
  geom_point(col=my_cols) +
  geom_text_repel(col=my_cols, size=2, max.overlaps=23)
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

```{r}
max <- which.max((candy[, "winpercent"] / candy[, "pricepercent"]))
candy[max,]
```

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head(candy[ord, c(11,12)], n=5)
```

Among the top 5 most expensive candy types, the least popular is Nik L Nip

## 5. Exploring the correlation structure

```{r}
#install.packages("corrplot")
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

chocolate and fruity

>Q23. Similarly, what two variables are most positively correlated?

chocolate and winpercent

## 6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
#pca$rotation[]
```
```{r}
plot(pca$x[,1], pca$x[,2])
```

```{r}
plot(pca$x[,1], pca$x[,2], col=my_cols, xlab="PC1", ylab="PC2", pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
head(my_data)
```

```{r}
p <- ggplot(my_data) + 
  aes(x=PC1, y=PC2,
      size=winpercent/100,
      text=rownames(my_data), 
      label=rownames(my_data)) +
  geom_point(col=my_cols)
p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 10)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
#install.packages("plotly")
#library(plotly)
#ggplotly(p)
```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard, and pluribus. Yes, they make sense because these types of candy usually go together (fruity candies are hard and pluribus) so it makes sense that they are all positive values. 